### 2023-07-19 python_excel_ite.tex
more citation cleaning

Spreadsheet based modeling has transformed the teaching of management science in business schools over the past twenty plus years. The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.  A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal. The next few decades saw numerous textbooks \cite{powellstepheng.BusinessAnalyticsArt, camm2020business} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.

@@

Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsoncatherineSoftwareEngineeringData2024,treadwayandrewSoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.  The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}. I hoped to address some of these issues in the EwP module. Sneakily, students would be exposed to some simple software engineering concepts during the process of building and analyzing familiar models. This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials. Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a bit more professional credibility.

@@

Now that we have created some useful modeling functionality, our goal is to make it easier for a modeler to actually use it. Certainly we do not want to rely on copying, pasting and modifying code when building a new model. In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezentomasWelcomePythonPackages} that can then be imported like any other Python library.

@@

A single Jupyter notebook is used to give a taste of using Python to automate the process of working with Excel files by tackling each of the four examples above. Each of these examples are based on a real problems I encountered either in research or industrial projects. Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \cite{moffittPracticalBusinessPython}. This blog, which focuses on helping Excel centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.

@@

The examples above barely scratch the surface of the use cases for Python integration with Excel. Our notebook ends with a shout out to the highly regarding Practical Business with Python blog and a list of specific examples that students can explore on their own.

@@

Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanucatalinMicrosoftConsidersAdding}. While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{cannystevePythonpptx2013}. Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate. Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar. Manipulation of MS Office documents is directly relevant to the job content of many of the students. This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.

>>>>

Spreadsheet based modeling has transformed the teaching of management science in business schools over the past twenty plus years. The late 1990's saw first editions of game changing textbooks \cite{winstonPracticalManagementScience2018,ragsdaleSpreadsheetModelingDecision2017} focused on the use of spreadsheets in teaching modeling and management science.  A series of conferences focused on teaching management science (TMS) with spreadsheets started at Dartmouth in 1998 and several presentations from that conference made up the bulk of papers in the first two issues of the INFORMS Transactions on Education journal. The next few decades saw numerous textbooks \cite{powellBusinessAnalyticsArt, camm2020business} and teaching papers \cite{INFORMSTransactionsEducation2014} that prominently featured spreadsheet based modeling.

@@

Recently there has been more recognition of the need for data scientists to learn basic software engineering principes \cite{nelsonSoftwareEngineeringData2024,treadwaySoftwareEngineeringData2023,rodriguesBuildingReproducibleAnalytical2023}.  The image of newly minted data science graduates developing spaghetti coded Jupyter notebooks not nearly ready for being put into production has reached \href{https://www.reddit.com/r/ProgrammerHumor/comments/y2xe3n/like_every_time_ever_when_the_devops_engineer/}{meme like status}. I hoped to address some of these issues in the EwP module. Sneakily, students would be exposed to some simple software engineering concepts during the process of building and analyzing familiar models. This is really no different than the integration of advanced Excel functions and techniques with spreadsheet modeling tutorials. Just as spreadsheet modeling students were able to easily wow their coworkers and managers with INDEX-MATCH and intermediate Pivot Table skills, I hoped that my students ability to properly document a Python function and use git and GitHub would give them a bit more professional credibility.

@@

Now that we have created some useful modeling functionality, our goal is to make it easier for a modeler to actually use it. Certainly we do not want to rely on copying, pasting and modifying code when building a new model. In the Python world, the natural next step is to deploy our code as a Python \textit{package} \cite{buezenWelcomePythonPackages} that can then be imported like any other Python library.

@@

A single Jupyter notebook is used to give a taste of using Python to automate the process of working with Excel files by tackling each of the four examples above. Each of these examples are based on a real problems I encountered either in research or industrial projects. Several web based resources are shared including the \href{https://pbpython.com/}{Practical Business Python blog} written by Chris Moffitt, a data analytics professional \cite{moffittPracticalBusinessPython2022}. This blog, which focuses on helping Excel centric analytics professionals transition some of their work to Python, includes numerous posts on different aspects of Python and Excel integration.

@@

The examples above barely scratch the surface of the use cases for Python integration with Excel. Our notebook ends with a shout out to the highly regarding \href{https://pbpython.com/}{Practical Business with Python} blog and a list of specific examples that students can explore on their own.

@@

Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{cimpanuMicrosoftConsidersAdding2017}. While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{cannyPythonpptx2013}. Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate. Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar. Manipulation of MS Office documents is directly relevant to the job content of many of the students. This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
